3D Magnetic modelling of ellipsoidal bodies

### 2017-03-03 manuscript/manuscript.tex

A considerable amount of literature has been published on the magnetic 
modelling of uniformly magnetized ellipsoids since the second half of 
the nineteenth century. Ellipsoids have flexibility to represent a wide 
range of geometrical forms, are the only known bodies which can be 
uniformly magnetized in the presence of a uniform inducing field and 
are the only bodies for which the self-demagnetization can be treated 
analytically. This property makes ellipsoids particularly useful for 
modelling compact orebodies having high susceptibility. In this case, 
neglecting the self-demagnetization may strongly mislead the interpretation 
of these bodies by using magnetic methods. A number of previous studies 
consider that the self-demagnetization can be neglected for the case in 
which the geological body has an isotropic susceptibility lower than or 
equal to 0.1 SI. This limiting value, however, seems to be determined 
empirically and there has been no discussion about how this value was 
determined. Besides, the geoscientific community lacks an easy-to-use 
tool to simulate the magnetic field produced by uniformly magnetized 
ellipsoids. Here, we present an integrated review of the magnetic 
modelling of arbitrarily oriented triaxial, prolate and oblate 
ellipsoids which have both induced and remanent magnetization, as 
well as isotropic or anisotropic susceptibility. We also propose a 
way of determining the isotropic susceptibility above which the 
self-demagnetization must be taken into consideration. Tests with 
synthetic data validate our approach. Finally, we provide a set 
of routines a set of routines to model the magnetic field produced 
by ellipsoids. The routines are written in Python language as 
part of the Fatiando a Terra, which is an open-source library 
for modelling and inversion in geophysics.

>>>>

A considerable amount of literature has been published on the magnetic 
modelling of uniformly magnetized ellipsoids since the second half of 
the nineteenth century. Ellipsoids have flexibility to represent a wide 
range of geometrical forms, are the only known bodies which can be 
uniformly magnetized in the presence of a uniform inducing field and 
are the only bodies for which the self-demagnetization can be treated 
analytically. This property makes ellipsoids particularly useful for 
modelling compact orebodies having high susceptibility. In this case, 
neglecting the self-demagnetization may strongly mislead the interpretation 
of these bodies by using magnetic methods. A number of previous studies 
consider that the self-demagnetization can be neglected for the case in 
which the geological body has an isotropic susceptibility lower than or 
equal to 0.1 SI. This limiting value, however, seems to be determined 
empirically and there has been no discussion about how this value was 
determined. Besides, the geoscientific community lacks an easy-to-use 
tool to simulate the magnetic field produced by uniformly magnetized 
ellipsoids. Here, we present an integrated review of the magnetic 
modelling of arbitrarily oriented triaxial, prolate and oblate 
ellipsoids. Our review includes ellipsoids with both induced and 
remanent magnetization, as well as with isotropic or anisotropic 
susceptibility. We also propose a 
way of determining the isotropic susceptibility above which the 
self-demagnetization must be taken into consideration. Tests with 
synthetic data validate our approach. Finally, we provide a set 
of routines to model the magnetic field produced 
by ellipsoids. The routines are written in Python language as 
part of the Fatiando a Terra, which is an open-source library 
for modelling and inversion in geophysics.
